Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets

### 2018-08-17 doc/Response.Rmd

The article "Adapting the Chumbley Score to match striae on Land Engraved Areass (LEAs) of bullets" extends an existing comparison algorithm for striated marks to the analysis of bullet LEAs. The paper makes two primary contributions. First, they modify the deterministic work of Hadler & Morris (2017) which itself is an extension of Chumbley et al. The authors' contribution is a modification of Hadler & Morris to reduce the number of scenarios which 'fail' under the Hadler & Morris approach. Second, they evaluate the work on two sets of bullet LEA data and explore some variants of the algorithm parameters to tune it for bullets rather than screwdrivers. Overall the paper presents a reasonable project to extend a nice algorithm for striated toolmark comparison.

Adaptation of the Chumbley el al algorithm from screwdrivers to bullets makes sense as both analyses involve striated marks. The authors here nicely point out that compared to screwdriver marks (as in the original Chumbley paper), LEAs are typically smaller both in length and width. LEAs also have a curved cross-section baseline which must be corrected.

The authors do a nice job covering previous work. I appreciate the summary tables 1 and 2. Table 1 is good as-is. Table 2 can be a bit misleading because we don't know what bullet datasets were used. Because these different papers analyzed different bullet data sets the underlying difficulty of matching impacts the performance. That is, bullets from very well marking firearms are 'easier' to match than those from firearms which mark poorly. So the FP and FN reported should be taken with a grain of salt. The authors should add a statement regarding different performance on different datasets.

*Thank you for pointing this out. You are right, we have added more information about the data, wherever available. We are also citing xxx to emphasize the difficulty level and resulting differences in error rates of different brands of firearms and ammunition*

@@

I'm also confused how the FP rate for Hare CNMS is 6.26% whereas all other FP are 6.25% This seems to likely be a typo. 6.25% imply close to 10,000 comparisons which I don't think the datasets have. I could be wrong but I wanted to point out this potential typo that should be checked.

*Thank you for spotting the typo. With respect to the sample size, the data set being used by \cite{aoas}. has about 10384 land-to-land comparisons and the error rates being reported here are computed from Receiver operating characteristic (ROC) curves. We decided to report False Negative error rates for a fixed value of False Positives (FP) = 6.25% to make them comparable to the observed values in Figure 10 (bottom right)*

@@

The authors describe the two methods of 'failure' for the method (Sec 2.3). I agree with the authors statement that this failure 'should' only happen for non-matches. It could also happen for matches if the optimal profiles are incorrect. That is, if there are two possible optimal match pairs (with very similar correlation scores), where the top scoring pair is incorrect but the second best scoring pair is correct. I'm not sure how often that happens. Overall, I would think that failing for non-matches would be ok. It's not clear that the described failures are the biggest cause of performance hits but improving the algorithm to handle these failures seems reasonable.

*You are right, a failure might occur for a match if the initial optimization fails. The new method is still susceptible to that. It would be interesting to further investigate thus problem in the furture. We have made a note of that in the conclusions*

@@

- p3 "Methods for matching marks for a variety of tools" should be changed to "Methods for matching striated marks for a variety of tools". This is an important point in that it helps make clear that the authors are studying and comparing striated toolmarks and not impressed toolmarks. I would also propose that the authors make this distinction earlier in the introduction.
- p3 "Analysis of these digitized markings require the use of statistical methods". Analysis does not *require* statistical methods. Useful information regarding common origin can be obtained from non-statistical methods. This being said, statistical methods are of course great and we should be working towards them. Authors could change to analysis "benefits from the use of statistical methods" or something similar.
- p14 how are the false negative and false positive rates computed?

Major Changes Required before acceptance:

- p8 "This severe limitation in the amount of available data poses the main challenge in". It's unclear what 'this' refers to. Do the authors mean the number of bullets and LEAs in the dataset or do they mean the number of sample points per striation profile (due to the fact that LEAs are shorter than screwdriver profiles)? Both would seem to be limitations but just because the profile is shorter doesn't inherently mean that there are fewer or less informative striaes. You could have a shorter profile with more informative features.

*We were indeed referring to the shorter length of the marks, which given the same amount of information decreases the power of the test, a purely technical consequence. Obviously you are right that this power calculation changes with a change in the amount of information*

@@

- Eq 1 uses a variable i, which is defined above as having the values of 1 or 2. I therefore don't understand how i<0 as written in Eq 1. This should be improved. Possibly this is a sloppy reuse of variables?

*thank you for pointing this out - we accidentally changed notations. this is fixed now. Index k=1,2 refers to the two different signatures, index i in Z refers to window locations.*

- How are the original profiles (eg. Figure 2 Left) identified from the 3D surface (eg. Figure 1 Top). That is, does the user simply draw a line from one end of the LEA to the other? How is the original 3D surface resampled to the 1d profile? This is very important because if you 'cut' the 3D profile the wrong way (at the wrong angle) then you will dilate or compress the extracted profile. This needs a lot more description.

*We followed the \cite{aoas} paper in extracting the profiles from scans. We have expanded on the discussion at the begining of section titled "Scans for land engraved areas" in paper*

@@

- Section 2.4 is not clearly written. I'm not confident that I understand their new method. This section should be rewritten to explain what the authors are doing. My guess is that rather than taking windows in integer number of steps away from the optimal position matching a left shift on one profile with a right shift on the other, the authors simply walk from the right-end to the left-end of one profile and from the left-end to the right-end of the other profile being careful to not compare same shift windows. Note that my description is not clear either as this is a tricky thing to put into words, but is something the authors need to do.

*Thank you for poining this out, you are right in your understanding of the method. We have included a figure for this and expanded out discussion, which hopefully makes the it easier to understand the method.*

@@

- The results presented on page 14 list the correct identification of "same-source and difference-source toolmarks of 93.5 to 94.1%". Does this mean the true positives are 93.5% and the true negatives are 94.1%? I don't think so because the next sentence lists the false negative rate of between 0.28 and 0.36. The authors need to explicitly list the true positive rate and the true negative rate. These must be separate because there are many more true negatives than true positives in any matching dataset. That means that an algorithm that always predicts "different source" would be correct most of the time. In fact, for the author's dataset the "always predict different" algorithm would be correct approximately 94% of the time (35 bullets where each has appx 2 matches). The lack of a true positive rate and true negative rate is is a major point that must be corrected. To be clear, the true positive rate, true negative rate, false positive rate, and false negative rate should all be reported for the first dataset, perhaps added to Table 3 or another table like Table 3.

*Thank you for bringing this to our notice. We have included a table of error rates. The 93.5% to 94.1% are the True Negative rates which are being mentioned. It is computed as 1 - Observed FPR. The observed False positive rates can be seen in figure 10. The True positive rates would then be 1 - Observed FNR or about 64% to 72%.*

@@

- The true positive rate, true negative rate, false positive rate, and false negative rate should also be presented for the second dataset (Hamby 44 scans, Sec 3.5). The authors do present AUC and EER curves but these are less informative and can often be used to hide weaker performing models.
- The results should all be listed as either fractions or all as percent (which is preferred) and not mix them.
- p21 the authors state "this kind of assessment is only feasible in the setting of a large study, such as the one we presented".  I strongly disagree with the use of the word "large". The presented work is not a large study. Ten firearms were used in each of two studies (it's not clear if both datasets used the same 10 firearms). A large study would include at least 100 different firearms and even 100 is not likely to provide statistical confidence. This is a very important point. If the paper is allowed to state that the described work is a large study it will set a precedent that will be detrimental to the field.

*We agree the data available to us is by no means large. Our comment was referring to the availability of 35 bullets rater than just 2 0r 3 as might be seen in actual case work*

@@

*By failed test, what is being meant is that the method fails to conduct a statistical test although there is a resolution available in the ground truth as whether the lands match or dont match. This is because of incorrect identification of Maximum correlations windows required for comparison of the two striation marks. This means that there is no resolution as to the two striations being a match or a non-match, when there should be one *

*Thank you for pointing this out. For the terminology error rates, we do mean FPR = FP/ (TN + FP) and FNR = FN/ (TP + FN), which holds with the definition of error rates in literature.*

*Thank you for pointing out the confusing terminology as seen in the "Missed identifications" Page 17. missed same-source marking. *

*We are aware of the current standards of classification as having a section as inconclusive for examination. But for the data sets we are working on, each and every land has been identified and none of the data being used, cannot be classified as from the same-source and different source*
